---
title: "GOV50 Final Project Thoughts"
author: "Elizabeth Esteban"
description: "Analyzing Members of Congress and their Constituents"
output:
  distill::distill_article:
    self_contained: false
---


## Project thoughts

For the final project, I'd like to analyze how effective Congress is. Members of Congress take on a big role of representing their constituents. Nevertheless, how well do they actually represent the people? Does party and interest groups take on a big role in their decision making? Do constituents solely vote on party preferences or are their any other factors that come in to play? In this project, I'll specifically be looking at the 116th House of Representatives. 


##Finding a data source:

Do our members of Congress do a good job on representing the people? In this study, I plan to examine the shared characteristics that lead to constituents electing an individual for Congress. Is this based on party preferences, or certain characterists? Does a member's committee reflect the needs of their district? 

Party or Constituents? What matters most for Members of Congress? Given how often members are divided into "yeas" and "nays" along the party line, it may seem evident that parties matter. Oftentimes, when casting for ostensibly party votes, members of Congress vote with fellow party members in spite of their disagreement of the policy or is it due to their agreement? (Krehbiel 1993). In this case study, I plan to measure district ideology using presidential vote share in districts. I'll be looking into each district's avergae Republican vote share from the 2012, 2016, and 2020 Presidential elections. The dependent variable in my analysis will be each member's DW-Nominate score and the explanatory variable is each district's ideology and the member's political party. 

##1st Data Visualization

```{r}
library(haven)
library(tidyr)
library(tidyverse)
congress <- read_dta("congress.dta")
```

```{r}
congress <- congress |>
  mutate(
    republican = if_else(party_code == 200,
                         "1","0"))
congress

congress <- congress|>
  mutate(
    Rpres_1220 = ((Romney12)/(Romney12 + Obama12) + (Trump16)/(Trump16 + Clinton16) + (Trump20)/(Biden20 + Trump20))/3)

congress

#scatterplot
scatterplot01 <- ggplot(congress) +
  geom_point(aes(x=Rpres_1220, y=nominate_dim1)) +
  labs(title="1st DW-Nominate by Republican Presidential Vote Share ",
       x="Republican Presidential Vote Share",
       y="First Dimension DW-Nominate")

ggsave(plot=scatterplot01, 
       filename = "Rpres_1220_scatter.pdf",
       width=7,
       height=5)

scatterplot01
```



