---
title: "GOV50 Final Project Thoughts"
author: "Elizabeth Esteban"
description: "Analyzing Members of Congress and their Constituents"
output:
  distill::distill_article:
    self_contained: false
---


## Project thoughts

For the final project, I'd like to analyze how effective Congress is. Members of Congress take on a big role of representing their constituents. Nevertheless, how well do they actually represent the people? Does party and interest groups take on a big role in their decision making? Do constituents solely vote on party preferences or are their any other factors that come in to play? In this project, I'll specifically be looking at the 116th House of Representatives. 


##Finding a data source:

Party or Constituents? What matters most for Members of Congress? Given how often members are divided into "yeas" and "nays" along the party line, it may seem evident that parties matter. Oftentimes, when casting for ostensibly party votes, members of Congress vote with fellow party members in spite of their disagreement of the policy or is it due to their agreement? (Krehbiel 1993). In this case study, I plan to measure district ideology using presidential vote share in districts. I'll be looking into each district's avergae Republican vote share from the 2012, 2016, and 2020 Presidential elections. The dependent variable in my analysis will be each member's DW-Nominate score and the explanatory variable is each district's ideology and the member's political party. 

##1st Data Visualization

```{r}
library(haven)
library(tidyr)
library(tidyverse)
congress <- read_dta("congress.dta")
```

```{r}
congress <- congress |>
  mutate(
    republican = if_else(party_code == 200,
                         "1","0"))
congress

congress <- congress|>
  mutate(
    Rpres_1220 = ((Romney12)/(Romney12 + Obama12) + (Trump16)/(Trump16 + Clinton16) + (Trump20)/(Biden20 + Trump20))/3)

congress

#scatterplot
scatterplot01 <- ggplot(congress) +
  geom_point(aes(x=Rpres_1220, y=nominate_dim1)) +
  labs(title="1st DW-Nominate by Republican Presidential Vote Share ",
       x="Republican Presidential Vote Share",
       y="First Dimension DW-Nominate")

ggsave(plot=scatterplot01, 
       filename = "Rpres_1220_scatter.pdf",
       width=7,
       height=5)

scatterplot01
```
#Scatterplot with regression line
```{r}
mod01 <- lm(data=congress, nominate_dim1 ~ Rpres_1220)

summary(mod01)
```

```{r}
mod01 <- lm(data=congress, nominate_dim1 ~ Rpres_1220)

scatterplot01_lm <- ggplot(congress) +
  geom_point(aes(x= Rpres_1220, y= nominate_dim1)) +
  labs(title="1st DW-Nominate by Republican Presidential Vote Share",
       x="Republican Presidential Vote Share",
       y="First Dimension DW-Nominate") +
  geom_abline(slope=mod01$coefficients[2], intercept = mod01$coefficients[1], 
              color="blue",
              linetype="dashed")
  
ggsave(plot=scatterplot01_lm, 
       filename = "Rpres_1220_scatter_lm.pdf",
       width=7,
       height=5)
scatterplot01_lm
```

#Basic scatterplot with labels and color-coding

```{r}
scatterplot02 <- ggplot(congress) +
  geom_text(aes(x=Rpres_1220, y=nominate_dim1, label= if_else(republican == 1, "R", "D"),
                color=factor(republican))) + 
  labs(title="1st DW-Nominate by Republican Vote Share",
       x="Republican Presidential Vote Share",
       y="First Dimension DW-Nominate") +
  scale_color_manual(name="Party", labels=c("Democrats", "Republicans"), values=c("Blue", "Red")) 

scatterplot02
```
#Scatterplot with regression lines for Democrats and Republicans
```{r}
mod02 <- lm(data=congress, nominate_dim1~ Rpres_1220 + republican)
summary(mod02)

scatterplot03 <- ggplot(congress) +
  geom_text(aes(x=Rpres_1220, y=nominate_dim1, label= if_else(republican == 1, "R", "D"),
                color=factor(republican))) + 
  labs(title="1st DW-Nominate by Republican Vote Share",
       x="Republican Presidential Vote Share",
       y="First Dimension DW-Nominate") +
  geom_abline(intercept= mod02$coefficients[1],
          slope = mod02$coefficients[2], 
              color="blue",
              linetype="dashed") +
  geom_abline(intercept= mod02$coefficients[1] + mod02$coefficients[3],
          slope = mod02$coefficients[2],
              color="blue",
              linetype="dashed") +
  scale_color_manual(name="Party", labels=c("Democrats", "Republicans"), values=c("Blue", "Red")) 

scatterplot03
```


The estimated coefficient on the Republican party indicator shows each party's mean and how the first dimension DW-Nominate changes given a one unit shift in the Republican Presidential vote share and the republican party indicator. Thus, this demonstrates how party's line with each Mcs' DW-Nominate scores(beliefs) and each district's party preference shown in the Republican Presidential vote Share. Thus, members of congress dw-nominate scores shows us how members do share similar attitudes as their constituents. (will expand more)

republican vote share = district ideology
